---
title: "ParkRNASeq"
format: html
editor: visual
---

## Loading data and preparing

```{r}
#| output: false
library("DESeq2")
library(DT)
```

```{r}
directory <- "htseq_out"
```

## DeSeq2 of all samples

Load metadata and output samples names

```{r}
sampleTable <- read.table("metadata.tsv", sep="\t", header = TRUE)
sampleTable$condition <- factor(sampleTable$Health.PD)
print(sampleTable$ID)
```

Load data and print number of transcripts

```{r}

ddsHTSeq <- DESeqDataSetFromHTSeqCount(sampleTable = sampleTable,
                                       directory = directory,
                                       design= ~ condition)

print(ddsHTSeq@elementMetadata@nrows)
```

Filter data to keep transcripts which have at least 10 reads in total

```{r}
keep <- rowSums(counts(ddsHTSeq)) >= 10
ddsHTSeq <- ddsHTSeq[keep,]
print(ddsHTSeq@elementMetadata@nrows)
```

Analysis

```{r}
dds <- DESeq(ddsHTSeq)

resultsNames(dds) 
res <- results(dds, name="condition_PD_vs_H")

```

```{r}
resOrderedAll <- res[order(res$pvalue),]

summary(resOrderedAll)

```

```{r}
datatable(as.data.frame(resOrderedAll@listData, row.names = resOrderedAll@rownames))
```

## DeSeq2 of SPb

Load metadata, filter SPb samples and output samples names

```{r}
sampleTable <- read.table("metadata.tsv", sep="\t", header = TRUE)
sampleTable$condition <- factor(sampleTable$Health.PD)
sampleTable <- sampleTable[sampleTable$Cells.Msc.SPb=='SPb', ]
print(sampleTable$ID)

```

Load data and print number of transcripts

```{r}

ddsHTSeq <- DESeqDataSetFromHTSeqCount(sampleTable = sampleTable,
                                       directory = directory,
                                       design= ~ condition)

print(ddsHTSeq@elementMetadata@nrows)
```

Filter data to keep transcripts which have at least 10 reads in total

```{r}
keep <- rowSums(counts(ddsHTSeq)) >= 10
ddsHTSeq <- ddsHTSeq[keep,]
print(ddsHTSeq@elementMetadata@nrows)
```

Analysis

```{r}
dds <- DESeq(ddsHTSeq)

resultsNames(dds) 
res <- results(dds, name="condition_PD_vs_H")

```

```{r}
resOrderedSPb <- res[order(res$pvalue),]

summary(resOrderedSPb)

```

```{r}
datatable(as.data.frame(resOrderedSPb@listData, row.names = resOrderedSPb@rownames))
```


## DeSeq2 of Msc

Load metadata, filter Msc samples and output samples names

```{r}
sampleTable <- read.table("metadata.tsv", sep="\t", header = TRUE)
sampleTable$condition <- factor(sampleTable$Health.PD)
sampleTable <- sampleTable[sampleTable$Cells.Msc.SPb=='Msc', ]
print(sampleTable$ID)

```

Load data and print number of transcripts

```{r}

ddsHTSeq <- DESeqDataSetFromHTSeqCount(sampleTable = sampleTable,
                                       directory = directory,
                                       design= ~ condition)

print(ddsHTSeq@elementMetadata@nrows)
```

Filter data to keep transcripts which have at least 10 reads in total

```{r}
keep <- rowSums(counts(ddsHTSeq)) >= 10
ddsHTSeq <- ddsHTSeq[keep,]
print(ddsHTSeq@elementMetadata@nrows)
```

Analysis

```{r}
dds <- DESeq(ddsHTSeq)

resultsNames(dds) 
res <- results(dds, name="condition_PD_vs_H")

```

```{r}
resOrderedMsc <- res[order(res$pvalue),]

summary(resOrderedMsc)

```

```{r}
datatable(as.data.frame(resOrderedMsc@listData, row.names = resOrderedMsc@rownames))
```


## Compare results

```{r}
print("All")
print(resOrderedAll@rownames[0:10])

print("SPb")
print(resOrderedSPb@rownames[0:10])

print("Msc")
print(resOrderedMsc@rownames[0:10])
```
## All transcripts with adj p-val <1e-10

```{r}
resOrderedAll <- na.omit(resOrderedAll)
resOrderedAllFiltered <- resOrderedAll[resOrderedAll$padj < 1e-10, ]

resOrderedAllFilteredNames <- resOrderedAllFiltered@rownames

resOrderedAllFilteredNames
```

## Msc transcripts with adj p-val <1e-10

```{r}
resOrderedMsc <- na.omit(resOrderedMsc)
resOrderedMscFiltered <- resOrderedMsc[resOrderedMsc$padj < 1e-10, ]

resOrderedMscFilteredNames <- resOrderedMscFiltered@rownames

resOrderedMscFilteredNames
```

## SPb transcripts with adj p-val <1e-10

```{r}
resOrderedSPb <- na.omit(resOrderedSPb)
resOrderedSPbFiltered <- resOrderedSPb[resOrderedSPb$padj < 1e-10, ]

resOrderedSPbFilteredNames <- resOrderedSPbFiltered@rownames

resOrderedSPbFilteredNames
```